### 2013-05-23 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
add bayesian version of Ricker
cleanup

```{r plotting-options}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)

library(knitcitations)

library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```
Load necessary libraries,

```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
library(nonparametricbayes) # loads the rest as dependencies
```

@@

Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r}
f <- RickerAllee
p <- c(2, 8, 5)
K <- p[2]
allee <- p[3]
```

@@

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 30  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
seed_i <- 111
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 35
```

@@

```{r obs, dependson="sdp-pars"}
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- 6
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
```

@@

```{r mle, dependson="obs"}
alt <- par_est(obs,  init = c(r = p[1], 
                              K = mean(obs$x[obs$x>0]), 
                              s = sigma_g))
est <- par_est_allee(obs, f, p,  
                     init = c(r = p[1] + 1, 
                              K = p[2] + 2, 
                              C = p[3] + 2, 
                              s = sigma_g))
```

@@

```{r}
gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
gp_dat <- gp_predict(gp, x_grid)

@@

```{r}
plots <- summary_gp_mcmc(gp)
plots[[1]]
plots[[2]]
```

```{r}
# Summarize the GP model
tgp_dat <- 
    data.frame(  x = x_grid, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
```

### Parametric Bayes

```{r}
# a bit unfair to start with the correct values, but anyhow...
init_p = p
names(init_p) = c("r0", "K", "theta")
y <- obs$x[-1] 
N=length(y);
```

We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.

```{r}
jags.data <- list("N","y")
n.chains = 1
n.iter = 40000
n.burnin = floor(10000)
n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.  The file defines the priors and the model, as seen when read in here

```{r}
cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
```

We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.  We use logarithms to maintain strictly positive values of parameters where appropriate.  Because our priors on the noise parameters are inverse gamma distributed.

```{r}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0","logtheta","stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(12345)

time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
)         
time_jags <- unname(time_jags["elapsed"]);
```

```{r}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")

# Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 0.01, 40)
logr_prior <- function(x) dunif(x, -6, 6)
logtheta_prior <- function(x) dunif(x, -6, 6)
stdQ_prior <- function(x) dunif(x, 0.001, 100)
stdR_prior <- function(x) dunif(x, 0.001, 100)

par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)

par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

# posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)

@@

```{r}
# um, cleaner if we were just be using the long form, par_posterior
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
colnames(mcmcall) <- who

pardist <- mcmcall
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])

bayes_coef <- apply(pardist,2,mean)
bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
bayes_pars
```

```{r}
par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
```

@@

```{r}

  true_means <- sapply(x_grid, f, 0, p)
  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
  est_means <- sapply(x_grid, est$f, 0, est$p)
  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)

models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
                     MLE=est_means, Ricker.MLE=alt_means, 
                     Parametric.Bayes = par_bayes_means)

@@

plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

@@

```{r}
MaxT = 1000
# uses expected values from GP, instead of integrating over posterior
# matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)

@@

Determine the optimal policy based on the true and MLE models

```{r}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

@@

matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

Determine the optimal policy based on parametric Bayesian model

```{r}
matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```

@@

```{r}
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

@@

```{r}

>>>>

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
library(knitcitations)
library(ggplot2) 
library(nonparametricbayes) 
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=12))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
posterior.mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

@@

Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r stateeq}
f <- RickerAllee
p <- c(1, 10, 5)
K <- 10  # approx, a li'l' less
allee <- 5 # approx, a li'l' less
```

@@

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 40
```

@@

```{r obs, dependson="sdp-pars"}
  set.seed(1234)
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- 2
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
raw_plot
```

@@

```{r mle, dependson="obs"}
set.seed(12345)
estf <- function(p){ 
    mu <- f(obs$x,0,p)
    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(p[1]+abs(rnorm(1,0,1)), 
         p[2]+rnorm(1,0,.1), 
         p[3]+rnorm(1,0, .1), 
         sigma_g + abs(rnorm(1,0,.1)))
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])

est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
```

Mean predictions

```{r mle-output, dependson="mle"}
true_means <- sapply(x_grid, f, 0, p)
est_means <- sapply(x_grid, est$f, 0, est$p)
```

@@

```{r gp, dependson=c("gp-priors", "obs")}
gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
```

```{r gp_traces_densities, dependson="gp"}
plots <- summary_gp_mcmc(gp)
```

```{r gp-output, dependson="gp"}
# Summarize the GP model
tgp_dat <- 
    data.frame(  x = x_grid, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
```

Parametric Bayesian Models
---------------------------

We use the JAGS Gibbs sampler, a recent open source BUGS
implementation with an R interface that works on most platforms.
We initialize the usual MCMC parameters; see `?jags` for details.

All parametric Bayesian estimates use the following basic parameters for the JAGS MCMC:

```{r jags-setup}
y <- obs$x[-1] 
N <- length(y);
jags.data <- list("N","y")
n.chains <- 3
n.iter <- 1e5
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

We will use the same priors for process and observation noise in each model,

```{r common-priors}
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(0.0001, 100)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
```

### Parametric Bayes of correct (Allen) model

```{r allen_pars}
init_p = c(2, K, 0.01)
names(init_p) = c("r0", "K", "theta")
```

The actual model is defined in a `model.file` that contains an R function
that is automatically translated into BUGS code by *R2WinBUGS*.  The file
defines the priors and the model. We write the file from R as follows:

```{r allen-model}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)

bugs.model <- 
paste(sprintf(
"model{
  K     ~ dunif(%s, %s)
  logr0    ~ dunif(%s, %s)
  logtheta ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  logr0_prior_p[1], logr0_prior_p[2],
  logtheta_prior_p[1], logtheta_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  theta <- exp(logtheta)
  x[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
    x[t+1] ~ dnorm(mu[t], iQ) 
  }
  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "allen.bugs")
```

Write the priors into a list for later reference

```{r allen-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, -6, 6)
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    logr0 = logr0_prior, logtheta = logtheta_prior,
                    stdQ = stdQ_prior, stdR = stdR_prior)
```

We define which parameters to keep track of, and set the initial values of
parameters in the transformed space used by the MCMC.  We use logarithms
to maintain strictly positive values of parameters where appropriate.

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here

# Can give a list of starting values or a function that can create random values 
jags.inits <- function(){
  list("K"=init_p["K"] * rlnorm(1,0,.5),
       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.5), 
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(12345)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
                model.file="allen.bugs", progress="none")

#Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
allen_jags <- autojags(allen_jags, n.update=10, n.iter=n.iter, 
                       n.thin = n.thin, progress="none") 
```

R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array. In this array each parameter is given its value at each sample from the posterior (index) for each chain.

```{r allen-traces, dependson="allen-mcmc"}
tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(allen_posteriors) = c("index", "variable", "value", "chain")
ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)
```

```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
allen_priors <- ddply(allen_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=3)
```

Reshape the posterior parameter distribution data, transform back into original space, and calculate the mean parameters and mean function

```{r allen-output, dependson="allen-traces"}
# A <- allen_posteriors
# A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
# pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])
bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
allen_means <- sapply(x_grid, f, 0, bayes_pars)
```

## Parametric Bayes based on the structurally wrong model (Ricker)

```{r ricker-model}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)

bugs.model <- 
paste(sprintf(
"model{
  K    ~ dunif(%s, %s)
  logr0    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  logr0_prior_p[1], logr0_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  x[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
    x[t+1] ~ dnorm(mu[t], iQ) 
  }
  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "ricker.bugs")
```

Compute prior curves

```{r ricker-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
```

We define which parameters to keep track of, and set the initial values of
parameters in the transformed space used by the MCMC.  We use logarithms
to maintain strictly positive values of parameters where appropriate.

```{r ricker-mcmc, dependson="ricker-model"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"] * rlnorm(1,0,.5),
       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                  model.file="ricker.bugs", progress.bar="none")
ricker_jags <- autojags(ricker_jags, n.update=10, n.iter=n.iter, 
                        n.thin = n.thin, progress.bar="none")
```

#### Convergence diagnostics for parametric bayes Ricker model

```{r ricker_traces, dependson="ricker-mcmc"}
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")

ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)
```

```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

Reshape posteriors data, transform back, calculate mode and corresponding function.

```{r ricker-output, dependson="ricker-traces"}
ricker_pardist <- acast(ricker_posteriors[2:3], 
                        1:table(ricker_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance")) 
ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
```

## Myers Parametric Bayes

```{r myers-model}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)

bugs.model <- 
paste(sprintf(
"model{
  logr0    ~ dunif(%s, %s)
  logtheta    ~ dunif(%s, %s)
  logK    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  logr0_prior_p[1], logr0_prior_p[2],
  logtheta_prior_p[1], logtheta_prior_p[2],
  logK_prior_p[1], logK_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);

  r0 <- exp(logr0)
  theta <- exp(logtheta)
  K <- exp(logK)

  x[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
    x[t+1] ~ dnorm(mu[t], iQ) 
  }

  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "myers.bugs")
```

```{r myers-priors}
logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
                    logr0 = logr_prior, logtheta = logtheta_prior, 
                    stdQ = stdQ_prior, stdR = stdR_prior)

@@

```{r myers-mcmc, dependson="myers-model"}
init_p = c(p[1], p[3], p[2])
names(init_p) = c("r0", "theta", "K")
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]) * rlnorm(1,0,.5), 
       "logK"=log(init_p["K"]),
       "logtheta" = log(init_p["theta"]) * rlnorm(1,0,.5),  
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
myers_jags <- autojags(myers_jags, n.update=10, n.iter=n.iter, 
                    n.thin = n.thin, progress.bar="none")
```

Convergence diagnostics for parametric bayes

```{r myers-traces, dependson="myers-mcmc"}
tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes the hard way...
myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(myers_posteriors) = c("index", "variable", "value", "chain")

ggplot(myers_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

@@

```{r myers-posteriors, dependson="myers-traces"}
par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=3)
```

```{r myers-output, dependson="myers-traces"}
myers_pardist <- acast(myers_posteriors[2:3], 
                        1:table(myers_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance")) 
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
```

@@

```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
                     MLE=est_means, Ricker=ricker_means, 
                     Allen = allen_means,
                     Myers = myers_means)

@@

model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
colorkey=cbPalette
names(colorkey) = model_names 
```

```{r Figure1, dependson="assemble-models"}
plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

## Step-ahead predictors

## Posterior predictive curves

This shows only the mean predictions.  For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.

We will need a vector version (`pmin` in place of `min`) of the parametric growth functions that can operate on the posteriors, (with appropriate ordering of parameters as they are in the posterior):

```{r par-fns}
ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])

```{r Figureb, dependson=c("assemble-models", "par-fns")}
require(MASS)
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
step_ahead_posteriors <- function(x){
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}

df_post <- step_ahead_posteriors(x)

ggplot(df_post) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
```

@@

```{r gp-opt, dependson="gp-output"}
MaxT = 1000
# uses expected values from GP, instead of integrating over posterior
#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)

@@

Determine the optimal policy based on the allen and MLE models

```{r mle-opt, dependson="mle-output"}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

@@

Determine the optimal policy based on Bayesian Allen model

```{r allen-opt, dependson="allen-output"}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

Bayesian Ricker

```{r ricker-opt, dependson="ricker-output"}
ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

Bayesian Myers model

```{r myers-opt, dependson="myers-output"}
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```

@@

```{r Figure2, dependson="assemble-opt"}
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

@@

```{r sims, dependson="assemble-opt"}
sims <- lapply(OPT, function(D){
  set.seed(1)
  lapply(1:100, function(i) 
    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
  )
})
